HTW Model

### 2024-02-22 Model/htw_model.qmd
cache big posterior dfs; re-render; create md renders

```{r}
#| cache: true

@@

future::plan(multisession)

@@

# organize test data predictions
post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
   })) |> 
  select(Fit_Method,pp,-data) |>  
  unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()

@@

# organize training data predictions
pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 20, return_dat="train_data")
   })) |>
  select(Fit_Method,pp,-data) |>
  unnest(pp) |> as.data.table() |> filter(expMode2=="Train")
plan(sequential)

@@

@@

```{r}
#| eval: true
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting testing data, aggregated over all participants and velocity bands. Note that Fit Method refers to how model parameters were optimized, while error values reflect mean absolute error for the 6 testing bands"
post_tabs <- abc_tables(post_dat,post_dat_l)
post_tabs$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() 
```

@@

Considering first the models fit to only the testing data, which reflect the best possible performance of ALM and EXAM at capturing the group-aggregated testing patterns. For the varied group, both ALM and EXAM are able to capture the median values of the observed data within the 66% credible intervals, and the spread of model predictions generally matches that of the observed data. For the constant group, only EXAM is able to capture the median range of values across the velocity bands, with ALM generally underestimating human velocoties in the upper bands, and overestimating in the lower bands. In the case of band 100, the median ALM prediction appears to match that of our participants - however this is due to a large subset of participants have ALM predictions near 0 for band 100, a pattern we will explore further in our considertation of individual patterns below.

@@

```{r}
#| label: fig-htw-post-dist
#| fig-cap: Posterior Distributions of $c$ and $lr$ parameters. Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. Note that the y axes of the plots for the c parameter are scaled logarithmically.
#| fig-height: 10
#| fig-width: 11

@@

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.4))
c_post + lr_post
```

@@

@@

post_tabs$agg_x_full |> flextable::tabulator(rows=c("Fit_Method","x"), columns=c("condit","Model"), 
                       `X` = as_paragraph(mean_error)) |> as_flextable()

@@

```{r fig.width=11, fig.height=10}
#| label: fig-htw-best-model
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 10
#| fig-width: 11
best_id_x <- post_dat_avg |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),abs_me=abs(mean_error)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

>>>>

```{r}
#| cache: false

@@

#future::plan(multisession)

@@

# organize test data predictions
# post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
#    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
#    })) |> 
#   select(Fit_Method,pp,-data) |>  
#   unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()
# 
# saveRDS(post_dat, here("data/model_cache/post_dat.rds"))

post_dat <- readRDS(here("data/model_cache/post_dat.rds"))

@@

# organize training data predictions
# pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
#    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 20, return_dat="train_data")
#    })) |>
#   select(Fit_Method,pp,-data) |>
#   unnest(pp) |> as.data.table() |> filter(expMode2=="Train")

#saveRDS(pd_train, here("data/model_cache/pd_train.rds"))

pd_train <- readRDS(here("data/model_cache/pd_train.rds"))

@@

#plan(sequential)

@@

```{r}
#| eval: true
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting testing data, aggregated over all participants and velocity bands. Note that Fit Method refers to how model parameters were optimized, while error values reflect mean absolute error for the 6 testing bands"
post_tabs <- abc_tables(post_dat,post_dat_l)

# post_tabs$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
#                        `ME` = as_paragraph(mean_error)) |> as_flextable()

post_tabs$agg_pred_full |> pander::pandoc.table()

@@

Considering first the models fit to only the testing data, which reflect the best possible performance of ALM and EXAM at capturing the group-aggregated testing patterns. For the varied group, both ALM and EXAM are able to capture the median values of the observed data within the 66% credible intervals, and the spread of model predictions generally matches that of the observed data. For the constant group, only EXAM is able to capture the median range of values across the velocity bands, with ALM generally underestimating human velocoties in the upper bands, and overestimating in the lower bands. In the case of band 100, the median ALM prediction appears to match that of our participants - however this is due to a large subset of participants have ALM predictions near 0 for band 100, a pattern we will explore further in our considertation of individual patterns below. Models fit to both training and testing data show a similar pattern to only the testing data display the same basic pattern as those fit to only the testing data, albeit with slightly larger residuals. However models fit to only the training data display markedly worse performance at accounting for the key testing patterns.

@@

```{r}
#| label: fig-htw-post-dist
#| fig-cap: Posterior Distributions of $c$ and $lr$ parameters. Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. Note that the y axes of the plots for the c parameter are scaled logarithmically.
#| fig-height: 9
#| fig-width: 11

@@

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.4)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.4))
c_post + lr_post
```

@@

```{r}
#| label: fig-htw-best-model
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 10
#| fig-width: 11

# could compute best model for each posterior parameter - examine consistency

tid1 <- post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
  mutate(e2=abs(y-pred)) |> 
  summarise(y1=mean(y), pred1=mean(pred),mean_error=abs(y1-pred1)) |>
  group_by(id,condit,Model,Fit_Method) |> 
  summarise(mean_error=mean(mean_error)) |> 
  arrange(id,condit,Fit_Method) |>
  round_tibble(1)

best_id <- tid1 |> 
  group_by(id,condit,Fit_Method) |> mutate(best=ifelse(mean_error==min(mean_error),1,0))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference<- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error) %>%
  pivot_wider(names_from = Model, values_from = c(mean_error)) %>%
  mutate(Error_difference = (ALM - EXAM))

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, Error_difference)) %>%
  ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
  geom_col() +
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")
```

@@

# post_tabs$agg_x_full |> flextable::tabulator(rows=c("Fit_Method","x"), columns=c("condit","Model"), 
#                        `X` = as_paragraph(mean_error)) |> as_flextable()

post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
  mutate(e2=abs(y-pred)) |> 
  summarise(y1=mean(y), pred1=mean(pred)) |>
  group_by(condit,Model,Fit_Method,x) |> 
  summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
  round_tibble(1) |> pander::pandoc.table()

# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
#   round_tibble(1) |> pander::pandoc.table()
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
#   select(-y,-pred) |>
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# 
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method,x) |> 
#   summarise(y2=mean(y1), pred2=mean(pred1)) |> 
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y2), pred=mean(pred2),mean_error=abs(y-pred)) |> 
#   select(-y,-pred) |>
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred),mean_error=abs(y1-pred1)) |>
#   group_by(id,condit,Model,Fit_Method) |> 
#   summarise(mean_error=mean(mean_error)) |> 
#   group_by(condit,Model,Fit_Method) |>
#   summarise(mean_error=mean(mean_error)) |> 
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()

@@

```{r fig.width=11, fig.height=10}
#| label: fig-htw-best-model-old
#| eval: false
#| include: false
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 10
#| fig-width: 11
best_id_x <- post_dat_avg |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),abs_me=abs(mean_error)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))
